---
title: "p8105_linear_models"
author: "Jana Lee"
date: "11/7/2019"
output: html_document
---

## Setup
```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
set.seed(1)
```


## Lecture: Linear Regression
lm = linear model, continuous outcome variable
glm = generalized linear models, binary outcome

arguments include:
- formula: y~x1 + x2
- data

Outputs are kind of a mess, but we can use the broom package to clean it up.

### Fit a model:
```{r}
library(p8105.datasets)

data("nyc_airbnb")

# Qustion: Is it possible to predict price based on a couple of preditors?
# Initial data cleaning:
nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(stars = review_scores_location / 2) %>% 
  rename(
    boro = neighbourhood_group,
    neighborhood = neighbourhood) %>% 
  filter(boro != "Staten Island") %>% 
  select(price, stars, boro, neighborhood, room_type)
```

Fit the first linear model:
When you fit 4 categorical variables, it is going to organize this by alphabetical order. Bronx becomes our reference category.
Stars = if stars go up by 1-unit, airbnb increase 31.99 dollars
Boro_Brooklyn = if brooklyn goes up about 1-unit, airbnb increases 40.50 dollars
```{r}
fit = lm(price ~ stars + boro, data = nyc_airbnb)

# this is kind of helpful: summary
summary(fit)

#Another way of extracting results
summary(fit)$coef

# extract coefficients out
coef(fit)

fitted.values(fit)

#tidying results, changing the string variable to make it easier to read, and presenting this in a table
fit %>% 
  broom::tidy() %>%
  mutate(term = str_replace(term, "boro", "Boro:")) %>% 
  knitr::kable(digits = 3)

```


Looking at factors...
```{r}
# put categorical factors in order based on what is most common, which borough is more frequent?
nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(
    boro = fct_infreq(boro),
    room_type = fct_infreq(room_type)
  )
```


Refitting the last model, now that factors have been reordered:
```{r}
fit = lm(price ~ stars + boro, data = nyc_airbnb)

fit %>% 
  broom::tidy()
```


Extract this information: r-squared, p-value, df, etc.
```{r}
fit %>% 
  broom::glance()
```


### Diagnostic information:
